---
title: "bedbaseRClient: illustrative operations on bedbase.org"
author: "Vincent J. Carey, stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{bedbaseRClient: illustrative operations on bedbase.org}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    number_sections: yes
---

```{r setup,echo=FALSE}
suppressMessages({
suppressPackageStartupMessages({
library(bedbaseRClient)
library(httr)
library(hca)
library(TFutils)
})
})
```

# Introduction

[bedbase.org]("https://bedbase.org") assembles genomic data on many
thousands of region-based scoring files.  See the [about page for bedbase.org](http://bedbase.org/about)
for details.

This package explores the resource using Bioconductor/R.  We'll use interesting
facilities of the hca package to explore JSON outputs of the bedbase API.

# A basic query

We'll find out about resources related to the GM12878 cell line.

## Metadata acquisition

The `get_bb_metadata` function is very simple.  We'll learn how
to vary the types of data retrieved later on.  One could question
whether 'GM12878' is a 'cell type' and we will examine this concept another time.

```{r lk1}
library(bedbaseRClient)
q1 = get_bb_metadata(query_type="cell_type", query_val="GM12878")
q1
cc = httr::content(q1)
names(cc[[1]][[2]])
```

## Metadata exploration

The content returned by the API is complex.  The hca package can help navigate.

```{r lk2}
library(hca)
lcc = lol(cc) # list of lists
lol_path(lcc)
```

With `lol_pull`, we can extract the content present along a given structural path.

```{r lk3}
exps = lol_pull(lcc, "[*][*].exp_protocol")
table(exps)
```

That's informative.  Let's check how many targets are transcription factors.
We'll use the Lambert table in the TFutils package.

```{r lk4}
library(TFutils)
lam = retrieve_lambert_main()
targs = lol_pull(lcc, "[*][*].target")
length(intersect(targs, lam$Name))
```

We can pivot to a different cell type as follows:
```{r lk5}
q2 = get_bb_metadata(query_type="cell_type", query_val="K562")
kk = httr::content(q2)
ktargs = lol_pull(lol(kk), "[*][*].target")
length(intersect(ktargs, lam$Name))
```
